Galaxy Zoo: Morphological Classifications for 150,000 Galaxies in HST Legacy Imaging

### 2016-05-12 writeup/gz_hubble_data.tex
Steven's suggested edits for section 1

% Note: 250-word limit
We present the data release paper for the Galaxy Zoo: Hubble (GZH) project.\footnote{This publication has been made possible by the participation of more than 200,000 volunteers in the Galaxy Zoo project. Their contributions are individually acknowledged at \url{http://authors.galaxyzoo.org/authors.html}.} This is the third phase in a large effort to measure reliable morphologies of galaxies on detailed scales by using crowdsourced visual classifications of color composite image. Images in GZH are selected from various \textit{Hubble Space Telescope} Legacy programs using the Advanced Camera for Surveys (AEGIS, COSMOS, GEMS, GOODS-N, and GOODS-S), with filters selected to probe the rest-frame optical emission from galaxies out to $z\simeq1$. The galaxies selected for GZH classifications go down to magnitude limits of $m_{I814W}<23.5$ and have a median redshift of $\langle z\rangle=0.9\pm0.6$, with a tail extending out to $z\simeq4$. The GZH morphological data includes measurements of both bulge- and disk-dominated galaxies, details on spiral disk structure that relate to the Hubble type, and numerous measurements of clump identification and geometry. This paper also describes and employs a novel method for calibrating morphological bias by using artificially-redshifted galaxy images as a baseline. The GZH catalog includes both the raw and calibrated morphological vote fractions for $150,771$ galaxies, providing a comprehensive data set suitable for large-scale studies of galaxy evolution.

@@

The morphology of galaxies encodes information on the orbital parameters and assembly history of the contents, including gas, dust, stars, and the central black hole. The morphology is also closely related to the local environment of the galaxy, as mutual interactions such as tides, shocks in cluster environments and direct mergers can all change the shape of the galaxy's gravitational potential. For $M^\star$~galaxies in the local Universe, this typically manifests at the most basic level as the difference between bulge-dominated, virialized systems resembling ellipticals (early-types) and disk-dominated, rotationally-supported disks (late-types) frequently exhibiting spiral arms. This dichotomy has been used to explore much of the astrophysics governing galaxy formation and evolution, and has been shown to be closely linked with other galactic properties such as stellar mass, halo mass, bolometric luminosity, black hole activity, effective radius, and the relative ages of the stellar populations.

The advent of larger telescopes in an increasing range of observing wavelengths has revealed that the distribution and properties of galaxy morphology have strongly evolved over the lifetime of the Universe. At redshift $z\simeq1$ (roughly 6~Gyr after the Big Bang), many galaxies are still in the process of assembling the baryonic mass required to reproduce the structures seen in the present day. This occurs in a variety of pathways, including accretion of baryons from large-scale galactic filaments onto halos via streaming, mergers of individual dark matter halos along with their baryons, conversion of gas into stars via gravitational collapse and star formation, etc. The process can also be slowed or even reversed via feedback from stellar winds, supernovae, and active black holes. Each of these processes affects the galaxy morphology in different ways, and so an accurate measurement of the demographics as a function of redshift provides an extremely powerful observational constraint on the physics involved \citep[for a recent review see][]{con14}.

Theoretical predictions for the morphology of galaxies as a function of redshift are primarily computed within the $\Lambda$CDM cosmological framework. Full treatments of this model gravitational interactions between baryons and dark matter, hydrodynamics of the gas, and baryonic physics related to star formation and evolution. The most advanced simulations now include volumes up to $\sim100$~Mpc$^3$ while simultaneously resolving the smaller scales necessary to reproduce the baryonic physics \citep{vog14a,sch15}. Such simulations predict clustering of galaxies on large scales in a hierarchical assembly model \citep{sil12}. The structure of individual galaxies is affected by the merger history \citep{too72,hop10}, local environment \citep[such as the morphology-density relation;][]{dre80}, initial dark halo mass, and many other factors. Morphologies of individual simulated high-mass galaxies at $z\sim2-3$ commonly show kpc-scale ``clumpy'' structures and few galaxies that are either smooth or well-ordered spirals; asymmetric profiles with strong density contrasts dominate populations in the early Universe until at least $z\sim1$ \citep{gen14}.

Observational studies of galaxies at high-redshift also display a wide range of morphological types, many of which are rare or absent at $z\sim0$. These include spheroids and disks (akin to the ellipticals and spirals seen in the local Universe), but also a significant population of massive, more irregular galaxies, including mergers, tadpoles, chains, double-clumps, and clump-clusters \citep{elm05,elm07,cam11a,for11a,kar15}. In contrast, while grand-design spirals have been observed as far back as $z=2.18$ \citep{law12}, their spatial density suggests that they are exceedingly rare, with a very low overall disk fraction \citep{mor13}. Current observational data thus strongly suggests that the classical Hubble sequence/tuning fork \citep{hub36} is not a suitable framework for characterizing high-redshift morphology.

Space-based observatories, particularly the \hubble, have been responsible for the bulk of imaging studies of high-redshift galaxies. Observations of fixed fields with very deep imaging \citep[eg,][]{wil96,gia04,bec06,dav07,sco07,gro11} give the photometric sensitivity necessary to detect $L^\star$ galaxies at $z>1$, while also providing the angular resolution to distinguish internal structure and characterize the morphology. While these measurements are helped by the fact that the angular diameter distance is relatively flat beyond $z>1$ in a flat $\Lambda$CDM cosmology, the relevant angular scales are only of the order $\sim5-10~\mathrm{kpc}/^{\prime\prime}$ \citep{wri06}. \hst{} can thus resolve much of the structure for a Milky Way-sized galaxy (at least for distinguishing a disk from a bulge), but will be limited for more compact structures. Since the size of galaxies evolves as roughly $r\propto(1+z)^{-1}$ \citep{law12a}, the compact sizes of high-redshift galaxies make detailed morphologies a challenge even for \hst{} \citep{che12}. The public availability of more than $10^5$~galaxies in archival imaging across various studies gives a data sample with the potential for high statistical significance.

One of the major difficulties in studying the morphologies of galaxies lies in the techniques used for measurement. Visual classification by experts has been used for many years \citep[eg,][]{hub26,dev59,san61,van76,nai10,bai11,kar15}. These methods have the advantage of using the significant processing power of the human brain to identify patterns, but suffer from issues such as lack of scaling to large surveys and potential issues with replicability and calibration. Automated measurements, both parametric \citep{pen02a,sim11,lac12} and non-parametric \citep{abr03,con03,lot04,sca07,bam08,fre13}, scale well to very large sample sizes, but do not always fully capture the relevant features, especially for asymmetric galaxies that become increasingly common at high redshifts. The Galaxy Zoo project \citep{lin08,lin11} utilizes crowdsourced visual classifications to measure galaxies in color-composite images. The efforts of more than $2\times10^5$~classifiers allow for multiple independent classifications of each image which are combined and calibrated to give a distribution of vote fractions proportional to the probability of a feature being visible. While crowdsourced data require extensive calibration \citep{bam09,wil13}, they have a proven reliability and have been used in dozens of papers \citep[eg,][]{lan08,bam09,dar10,mas11c,ski12,sim13,sch14,wil15}.

This paper presents the results from the Galaxy Zoo Hubble (GZH) project.\footnote{\url{http://zoo3.galaxyzoo.org/}} GZH was the third phase of Galaxy Zoo, following its initial results classifying $\sim900,000$~SDSS~images into primarily early/late~types \citep{lin11} and Galaxy~Zoo~2, which covered $\sim250,000$~SDSS~images using a more detailed classification scheme that included bars, spiral arms, and galactic bulges \citep{wil13}. GZH used a similarly detailed classification scheme, but focused for the first time on images of high-redshift galaxies taken with the \hubble. The Galaxy Zoo: CANDELS project has also classified morphologies of galaxies at high redshift using ACS and WFC3 imaging (Simmons et~al., submitted).

@@

This paper uses the WMAP9 cosmology parameters of $(\Omega_m,\Omega_\Lambda,h)=(0.258,0.718,0.697)$ \citep{hin13}.

@@

>>>>

% Note: 250-word limit
We present the data release paper for the Galaxy Zoo: Hubble (GZH) project.\footnote{This publication has been made possible by the participation of more than 200,000 volunteers in the Galaxy Zoo project. Their contributions are individually acknowledged at \url{http://authors.galaxyzoo.org/authors.html}.} This is the third phase in a large effort to measure reliable morphologies of galaxies on detailed scales by using crowdsourced visual classifications of color composite image. Images in GZH are selected from various \textit{Hubble Space Telescope} Legacy programs using the Advanced Camera for Surveys (AEGIS, COSMOS, GEMS, GOODS-N, and GOODS-S), with filters selected to probe the rest-frame optical emission from galaxies out to $z\simeq1$. The galaxies selected for GZH classifications go down to magnitude limits of $m_{I814W}<23.5$ and have a median redshift of $\langle z\rangle=0.9\pm0.6$, with a tail extending out to $z\simeq4$. The GZH morphological data includes measurements of both bulge- and disk-dominated galaxies, details on spiral disk structure that relate to the Hubble type, as well as clump identification and geometry. This paper also describes and employs a novel method for calibrating morphological bias by using artificially-redshifted galaxy images as a baseline. The GZH catalog includes both the raw and calibrated morphological vote fractions for $150,771$ galaxies, providing a comprehensive data set suitable for large-scale studies of galaxy evolution.

@@

The morphology of galaxies encodes information on the orbital parameters and assembly history of their contents, including gas, dust, stars, and the central black hole. The morphology is also closely related to the local environment of a galaxy, as interactions such as tides, shocks in cluster environments and direct mergers can all influence a galaxy's internal structure.  For $M^\star$~galaxies in the local Universe, this interplay between the physical development of a galaxy and its external appearance typically manifests at the most basic level as the difference between bulge-dominated, virialized systems resembling ellipticals (early-types) and disk-dominated, rotationally-supported systems (late-types) frequently exhibiting spiral arms. This dichotomy has been used to explore much of the astrophysics governing galaxy formation and evolution, and has been shown to be closely linked with other galactic properties such as stellar mass, halo mass, bolometric luminosity, black hole activity, effective radius, and the relative ages of the stellar populations.

The advent of larger telescopes in an increasing range of observing wavelengths has revealed that the distribution and properties of galaxy morphology have strongly evolved over the lifetime of the Universe. At redshift $z\simeq1$ (roughly 6~Gyr after the Big Bang), many galaxies are still in the process of assembling the baryonic and stellar mass required to reproduce the mature, coherent structures seen at the present day. This growth occurs via a variety of pathways, including accretion from large-scale galactic filaments onto halos via streaming, mergers of individual halos, conversion of gas into stars via gravitational collapse, etc. The accrual of baryons can be slowed or even reversed via feedback from stellar winds, supernovae, and active black holes. Each of these processes affects galaxy morphology in different ways, and so an accurate measurement of morphological demographics as a function of redshift provides an extremely powerful observational constraint on the physics involved \citep[for a recent review see][]{con14}.

Theoretical predictions for the morphology of galaxies as a function of redshift are primarily computed within the $\Lambda$CDM cosmological framework. Full treatments model gravitational interactions between baryons and dark matter, hydrodynamics of the gas, and baryonic physics related to star formation and evolution. The most advanced simulations now span volumes up to $\sim100$~Mpc$^3$ while simultaneously resolving the smaller ($<1$ kpc) scales necessary to reproduce the influence of baryonic physics \citep{vog14a,sch15}. Such simulations predict clustering of galaxies on large scales in a hierarchical assembly model \citep{sil12}. The structure of individual galaxies is affected by their merger history \citep{too72,hop10}, local environment \citep[such as the morphology-density relation;][]{dre80}, initial dark halo mass, and many other factors. Morphologies of individual simulated high-mass galaxies at $z\sim2-3$ commonly show kpc-scale ``clumpy'' structures, with few galaxies that are either smooth or well-ordered spirals; asymmetric profiles with strong density contrasts dominate simulated populations in the early Universe until at least $z\sim1$ \citep{gen14}.

Observational studies of galaxies at high-redshift also display a wide range of morphological types, many of which are rare or absent at $z\sim0$. Although spheroids and disks are present, they are typically much more compact than those in the local Universe. There is also a significant population of more irregular, massive galaxies, including mergers, tadpoles, chains, double-clumps, and clump-clusters \citep{elm05,elm07,cam11a,for11a,kar15}. In contrast, while grand-design spirals have been observed as far back as $z=2.18$ \citep{law12}, their spatial density suggests that they are exceedingly rare, with a very low fraction of undisturbed disks \citep{mor13}. Current observational data thus strongly suggests that the classical Hubble sequence/tuning fork \citep{hub36} is not a suitable framework for characterizing high-redshift morphology.

Space-based observatories, particularly the \hubble (\hst), have been responsible for the bulk of imaging studies of high-redshift galaxies. Observations of fixed fields with very deep imaging \citep[eg,][]{wil96,gia04,bec06,dav07,sco07,gro11} give the photometric sensitivity necessary to detect $L^\star$ galaxies at $z>1$, while also providing the angular resolution to distinguish internal structure and characterize the morphology. While these measurements are helped by the fact that the angular diameter distance is relatively flat beyond $z>1$ in a flat $\Lambda$CDM cosmology, the relevant angular scales are only of the order $\sim5-10~\mathrm{kpc}/^{\prime\prime}$ \citep{wri06}. \hst{} can thus resolve much of the structure for a Milky Way-sized galaxy (at least for distinguishing a disk from a bulge), but will be limited for more compact structures. Since the size of galaxies evolves as roughly $r\propto(1+z)^{-1}$ \citep{law12a}, the compact sizes of high-redshift galaxies make detailed morphologies a challenge even for \hst{} \citep{che12}. The public availability of more than $10^5$~galaxies in archival imaging across various studies gives a data sample with the potential for high statistical significance.

One of the major difficulties in studying the morphologies of galaxies lies in the techniques used for measurement. Visual classification by experts has been used for many decades \citep[eg,][]{hub26,dev59,san61,van76,nai10,bai11,kar15}. These methods have the advantage of using the significant processing power of the human brain to identify patterns, but suffer from issues such as lack of scaling to large surveys and potential issues with replicability and calibration. Automated measurements, both parametric \citep{pen02a,sim11,lac12} and non-parametric \citep{abr03,con03,lot04,sca07,bam08,fre13}, scale well to very large sample sizes, but do not always fully capture the relevant features, especially for asymmetric galaxies that become increasingly common at high redshifts. The Galaxy Zoo project \citep{lin08,lin11} utilizes crowdsourced visual classifications to measure galaxies in color-composite images. The efforts of more than $2\times10^5$~classifiers allow for multiple independent classifications of each image which are combined and calibrated to give a distribution of vote fractions proportional to the probability of a feature being visible. While crowdsourced data require extensive calibration \citep{bam09,wil13}, they have a proven reliability and have been used in dozens of papers \citep[eg,][]{lan08,bam09,dar10,mas11c,ski12,sim13,sch14,wil15}.

This paper presents the results from the Galaxy Zoo Hubble (GZH) project.\footnote{\url{http://zoo3.galaxyzoo.org/}} GZH was the third phase of Galaxy Zoo, following its initial results classifying $\sim900,000$~SDSS~images into primarily early/late~types \citep{lin11} and Galaxy~Zoo~2, which covered $\sim250,000$~SDSS~images using a more detailed classification scheme that included bars, spiral arms, and galactic bulges \citep{wil13}. GZH used a similarly detailed classification scheme, but focused for the first time on images of high-redshift galaxies taken with \hst. The Galaxy Zoo: CANDELS project has also classified morphologies of galaxies at high redshift using ACS and WFC3 imaging (Simmons et~al., submitted).

@@

This paper assumes the WMAP9 cosmological parameters of $(\Omega_m,\Omega_\Lambda,h)=(0.282,0.718,0.697)$ \citep{hin13}.

@@

KS gratefully acknowledges support from Swiss National Science Foundation Grant PP00P2\_138979/1.
